Hierarchical Reverberation Mapping

### 2013-12-03 Paper/ms.tex
a --> k and other tweaks

\section{Combining Inferences About Multiple Objects}\label{sec:multiple}
Consider a sample of $N$ objects, each of which has parameters $\theta_i$.
If the $i$th object is analysed on its own, the inference about 
its parameters $\theta_i$ is described by a posterior distribution
\begin{eqnarray}
p(\theta_i | x_i) \propto \pi(\theta_i)p(x_i | \theta_i)\label{eq:individual}
\end{eqnarray}
where $\pi(\theta_i)$ is the prior distribution and $p(x_i | \theta_i)$ is the
likelihood function. If $N$ objects are analysed separately, inferences can
be made about the diversity of $\theta_i$ across the sample. However, these
inferences may be incorrect due to the implicit assumption that the prior for
all of the $\btheta$ is independent \citep{2013arXiv1310.5177B}.
A hierarchical model may be used to overcome this problem.

@@

The posterior distribution for the lag $\bar{\tau}$ of a single object
$i$ may be obtained
by fitting the following model. We assume, for simplicity, that
the transfer function is uniform
between limits $a$ and $b$, where $b > a$:
\begin{eqnarray}
\Psi(\tau) &=& \left\{
\begin{array}{lr}
\frac{1}{b-a}, & \tau \in [a,b]\\
0, & \textnormal{otherwise}
\end{array}\right.
\end{eqnarray}
and our goal is to measure the mean lag:
\begin{eqnarray}
\bar{\tau} &=& \int \tau \Psi(\tau) \, d\tau\\
&=& \frac{1}{2}(b-a).
\end{eqnarray}
Note that inferring $\bar{\tau}$ from the data requires that we
marginalise over an infinite number of nuisance parameters describing the
behaviour of $y(t)$ at unobserved times \citep{pancoast}.
The prior for the underlying time variation of the continuum emission is
a continuous autoregressive process of order 1, or a CAR(1) model. These models
have been studied extensively for AGN variability
\citep[e.g.][]{2009ApJ...698..895K, 2011ApJ...735...80Z, 2013ApJ...765..106Z}.
This marginalisation can be done
either analytically or inside MCMC. We used the latter approach for simplicity.
We discretised
time using ten time bins per day, and have a discrete continuum light curve
$\mathbf{y} = \{y_1, ..., y_n\}$, included in the model as a set of unknown
parameters. The prior for these parameters is:
\begin{eqnarray}
p(y_i | y_{i-1},  m, a, \beta) \sim \mathcal{N}
\left(m + a\left(y_{i-1} - m\right), \beta^2\right)
\end{eqnarray}
for $i \geq 2$. This is the discrete AR(1) model from time series theory.
$m$ describes the mean level of the continuum light curve, $\beta$ controls
the size of the short-term fluctuations, and $a$ controls the correlation
timescale (the timescale is given by $L = -1/\ln(a)$).

For the likelihood (or sampling distribution, really the prior for the data
given the parameters) we made the conventional gaussian assumption:
\begin{eqnarray}
Y_i &\sim& \mathcal{N}\left(y(t_{y_i}), \sigma_{y_i}^2\right)\\
L_i &\sim& \mathcal{N}\left(l(t_{l_i}), \sigma_{l_i}^2\right)
\end{eqnarray}

@@

\section{Demonstration on Simulated Data}\label{sec:data}
To test our hierarchical model, and compare it to the stacked cross-correlation
function, we simulated data from a sample of 100 AGN. We assumed the
continuum flux for each object fluctuates around a mean value $m = 50$ in
arbitrary units. The variability timescales $L$ were simulated from a very
broad log-uniform distribution (i.e. a uniform distribution for $\ln(L)$)
with a lower limit of 10 days and an upper limit of $1,000$ days. The
variability amplitudes $\beta$ were also simulated from a log-uniform
distribution with lower limit of 0.2 flux units and an upper limit of 1 flux
unit. This implies some objects are more variable than others, as is typical
in reverberation mapping.

Each object in the sample had its own transfer function parameters. The upper
limit $b$ of the transfer function was simulated using a log-normal distribution
with a median value of 10 days, and the standard deviation of $\ln(b)$ was
0.3. Given $b$, $a$ was simulated from a uniform distribution between $0$ and
$b$. The resulting distribution of lags $\bar{\tau} = (a+b)/2$ is very well
approximated by a lognormal distribution with a median of 7.4 days and width
(quantified by the standard deviation of $\log_{10}(\bar{\tau})$) of 0.157
(these two quantities are the ones we will infer using the hierarchical model).
The response amplitudes $A$ and $C$ were set to 0.5 and 0 respectively, for
all objects.

@@

Finally, we combined the results from the single-object model into an inference
about hyperparameters $\mu$ and $\sigma$ describing the central value and
the diversity of $\bar{\tau}$ respectively, using the result of
Equation~\ref{eq:individual}. This requires us to know the prior $\pi(\bar{\tau})$
implied by the simple-object model, since we did not assign the prior directly
to $\bar{\tau}$. The prior for $\bar{\tau}$ in the single-object model is
very well approximated by a log-uniform distribution.

@@

\section{Conclusions}\label{sec:conclusions}
Reverberation mapping campaigns are observationally intensive, yet potentially
very informative. This has led to
research into RM data analysis techniques, as well as less costly observing
strategies. The recent approach of \citet{2012MNRAS.427.2701F} demonstrated
that it is possible to combine information from many sparsely-measured AGN
and infer properties of the sample rather than insisting on accurate results
about a particular object.

@@

\section*{Acknowledgements}
It is a pleasure to thank Tommaso Treu, Anna Pancoast (UCSB), and
Brandon Kelly (UCSB) for many
useful conversations about reverberation mapping. The STAN ({\tt mc-stan.org})
team provided helpful advice about the use of their software. BJB is partially
supported by the Marsden Fund (Royal Society of New Zealand). We are grateful
to the referee, Stephen Fine, for his helpful comments.

>>>>

\vspace{-0.5cm}
\section{Combining Inferences About Multiple Objects}\label{sec:multiple}
Consider a sample of $N$ objects, each of which has parameters $\theta_i$.
If the $i$th object is analysed on its own, the inference about 
its parameters $\theta_i$ is described by a posterior distribution
\begin{eqnarray}
p(\theta_i | x_i) \propto \pi(\theta_i)p(x_i | \theta_i)\label{eq:individual}
\end{eqnarray}
where $\pi(\theta_i)$ is the prior distribution and $p(x_i | \theta_i)$ is the
likelihood function. If $N$ objects are analysed separately, inferences can
be made about the diversity of $\theta_i$ across the sample. However, these
inferences may be incorrect due to the implicit assumption that the prior for
all of the $\btheta$ is independent \citep{2013arXiv1310.5177B}.
A hierarchical model may be used to overcome this problem.

@@

The posterior distribution for the lag $\bar{\tau}$ of a single object
$i$ may be obtained
by fitting the following model. We assume, for simplicity, that
the transfer function is uniform
between limits $a$ and $b$, where $b > a$:
\begin{eqnarray}
\Psi(\tau) &=& \left\{
\begin{array}{lr}
\frac{1}{b-a}, & \tau \in [a,b]\\
0, & \textnormal{otherwise}
\end{array}\right.
\end{eqnarray}
and our goal is to measure the mean lag:
\begin{eqnarray}
\bar{\tau} &=& \int \tau \Psi(\tau) \, d\tau\\
&=& \frac{1}{2}(b-a).
\end{eqnarray}
Note that inferring $\bar{\tau}$ from the data requires that we
marginalise over an infinite number of nuisance parameters describing the
behaviour of $y(t)$ at unobserved times \citep{pancoast}.
The prior for the underlying time variation of the continuum emission is
a continuous autoregressive process of order 1, or a CAR(1) model. These models
have been studied extensively for AGN variability
\citep[e.g.][]{2009ApJ...698..895K, 2011ApJ...735...80Z, 2013ApJ...765..106Z}.
This marginalisation can be done
either analytically or inside MCMC. We used the latter approach for simplicity.
We discretised
time using ten time bins per day, and have a discrete continuum light curve
$\mathbf{y} = \{y_1, ..., y_n\}$, included in the model as a set of unknown
parameters. The prior for these parameters is:
\begin{eqnarray}
p(y_i | y_{i-1},  m, k, \beta) \sim \mathcal{N}
\left(m + k\left(y_{i-1} - m\right), \beta^2\right)
\end{eqnarray}
for $i \geq 2$. This is the discrete AR(1) model from time series theory.
$m$ describes the mean level of the continuum light curve, $\beta$ controls
the size of the short-term fluctuations, and $k$ controls the correlation
timescale (the timescale itself is given by $L = -1/\ln(k)$).

For the likelihood (or sampling distribution, really the prior for the data
given the parameters) we made the conventional assumption of ``Gaussian noise'':
\begin{eqnarray}
Y_i &\sim& \mathcal{N}\left(y(t_{y_i}), \sigma_{y_i}^2\right)\\
L_i &\sim& \mathcal{N}\left(l(t_{l_i}), \sigma_{l_i}^2\right)
\end{eqnarray}

@@

\vspace{-0.5cm}
\section{Demonstration on Simulated Data}\label{sec:data}
To test our hierarchical model, and compare it to the stacked cross-correlation
function, we simulated data from a sample of 100 AGN. We assumed the
continuum flux for each object fluctuates around a mean value $m = 50$ in
arbitrary units. The variability timescales $L$ were simulated from a very
broad log-uniform distribution (i.e. a uniform distribution for $\ln(L)$)
with a lower limit of 10 days and an upper limit of $1,000$ days. The
variability amplitudes $\beta$ were also simulated from a log-uniform
distribution with lower limit of 0.2 flux units and an upper limit of 1 flux
unit. This implies some objects are more variable than others, as is typical
in reverberation mapping.

Each object in the sample had its own transfer function parameters. The upper
limit $b$ of the transfer function was simulated using a log-normal distribution
with a median value of 10 days, and the standard deviation of $\ln(b)$ was
0.3. Given $b$, $a$ was simulated from a uniform distribution between $0$ and
$b$. The resulting distribution of lags $\bar{\tau} = (a+b)/2$ is very well
approximated by a lognormal distribution with a median of 7.4 days and width
(quantified by the standard deviation of $\log_{10}(\bar{\tau})$) of 0.157
(these two quantities are the ones we will infer using the hierarchical model).
The response amplitudes $A$ and $C$ were set to 0.5 and 0 respectively, for
all objects. Note that the distribution of the parameters of the 100 objects is
not the same as the priors used when analysing them with the single-object model.

@@

Finally, we combined the results from the single-object model into an inference
about hyperparameters $\mu$ and $\sigma$ describing the central value and
the diversity of $\bar{\tau}$ respectively, using the result of
Equation~\ref{eq:individual}. This requires us to know the prior $\pi(\bar{\tau})$
implied by the simple-object model, since we did not assign the prior directly
to $\bar{\tau}$ but indirectly through $a$ and $b$.
The prior for $\bar{\tau}$ in the single-object model is
very well approximated by a log-uniform distribution.

@@

\vspace{-0.5cm}
\section{Conclusions}\label{sec:conclusions}
Reverberation mapping campaigns are observationally intensive, yet potentially
very informative. This has led to
research into RM data analysis techniques, as well as less costly observing
strategies. The recent approach of \citet{2012MNRAS.427.2701F} demonstrated
that it is possible to combine information from many sparsely-measured AGN
and infer properties of the sample rather than insisting on accurate results
about a particular object.

@@

\vspace{-0.5cm}
\section*{Acknowledgements}
It is a pleasure to thank Tommaso Treu, Anna Pancoast (UCSB), and
Brandon Kelly (UCSB) for many
useful conversations about reverberation mapping. The STAN ({\tt mc-stan.org})
team provided helpful advice about the use of their software. BJB is partially
supported by the Marsden Fund (Royal Society of New Zealand). We are grateful
to the referee, Stephen Fine, for his helpful comments.
